### 2025-05-28 docs/article/Sections/benchmark.tex
Section 4: Include approaches from Spring MVC and Quarkus

The benchmark is designed with a modular architecture, separating the
\textit{view} and \textit{model} layers from the \textit{controller} layer
~\cite{Bucanek2009}, which allows for easy extension and integration of new
template engines and frameworks. It also includes a set of tests to ensure the
correctness of implementations and to validate the \textit{HTML} output. It
includes two different data models, defined as \texttt{Presentation} and
\texttt{Stock}, shown in \autoref{lst:stock-class} and
\autoref{lst:presentation-class}, respectively. The \texttt{Presentation} class
represents a presentation with a title, speaker name, and summary, while the
\texttt{Stock} class represents a stock with a name, URL, symbol, price,
change, and ratio.

@@

The Spring WebFlux core implementation uses Project Reactor to support a
reactive programming model: each method returns a \texttt{Flux<String>} as the
response body, which acts as a publisher that progressively streams the HTML
content to the client. The implementation includes three main approaches to
PSSR\@:
\begin{itemize}
  \item \textbf{Reactive}: The template engine is used in a reactive context,
        where the HTML content rendered using the reactive programming model.
        An example of this approach is the Thymeleaf template engine
        when using the \texttt{ReactiveDataDriverContextVariable} in conjunction with a
        non-blocking Spring \texttt{ViewResolver}.
  \item \textbf{Virtual}: The template engine is used in a non-blocking
        context, where the HTML content is rendered within the context of Virtual
        Threads. This method is used for the template engines that do not traditionally
        support non-blocking I/O, be it either because they use external DSLs and in
        consequence only support the blocking \texttt{Iterable} interface, or
        because they do not support the asynchronous rendering of HTML content.
  \item \textbf{Suspendable}: The template engine is used in a suspendable
        context, where the HTML content is rendered within the context of a suspending function.
        An example of this approach is the HtmlFlow template engine, which
        supports suspendable templates with the use of the \texttt{Flow} class from the
        Kotlin standard library.
\end{itemize}
% It also includes methods using Kotlin coroutines and other
% asynchronous mechanisms supported by template engines, such as the
% \texttt{writeAsync} method from HtmlFlow, which enables non-blocking I/O
% using \textit{continuations}. 
The Spring MVC implementation uses handlers based solely on the blocking
interface of the \texttt{Observable} class. To enable PSSR in this context, we
utilize the \texttt{StreamingResponseBody} interface, which allows the
application to write directly to the response \texttt{OutputStream} without
blocking the servlet container thread. According to the Spring documentation,
this class is \textit{a controller method return value type for asynchronous
  request processing where the application can write directly to the response
  OutputStream without holding up the Servlet container thread}

@@

@@

>>>>

The benchmark\footnote{\url{https://github.com/fmcarvalho/comparing-non-blocking-progressive-ssr}}
is designed with a modular architecture, separating the \textit{view} and
\textit{model} layers from the \textit{controller} layer ~\cite{Bucanek2009},
which allows for easy extension and integration of new template engines and
frameworks. It also includes a set of tests to ensure the correctness of
implementations and to validate the \textit{HTML} output. It includes two
different data models, defined as \texttt{Presentation} and \texttt{Stock}, as
shown in \autoref{lst:stock-class} and \autoref{lst:presentation-class},
following the same model as
\textit{spring-webflux-comparing-template-engines}~\footnote{\url{https://github.com/xmlet/spring-webflux-comparing-template-engines}}.
The \texttt{Presentation} class represents a presentation with a title, speaker
name, and summary, while the \texttt{Stock} class represents a stock with a
name, URL, symbol, price, change, and ratio.

@@

The Spring WebFlux core implementation uses Project Reactor to support a
reactive programming model: each method returns a \texttt{Flux<String>} as the
response body, which acts as a publisher that progressively streams the HTML
content to the client. The implementation includes four main approaches to
PSSR\@:
\begin{itemize}
  \item \textbf{Reactive}: The template engine is used in a reactive context,
        where the HTML content rendered using the reactive programming model.
        An example of this approach is the Thymeleaf template engine
        when using the \texttt{ReactiveDataDriverContextVariable} in conjunction with a
        non-blocking Spring \texttt{ViewResolver}.
  \item \textbf{Virtual}: The template engine is used in a non-blocking
        context, where the HTML content is rendered within the context of Virtual
        Threads. This method is used for the template engines that do not traditionally
        support non-blocking I/O, be it either because they use external DSLs and in
        consequence only support the blocking \texttt{Iterable} interface, or
        because they do not support the asynchronous rendering of HTML content.
  \item \textbf{Suspendable}: The template engine is used in a suspendable
        context, where the HTML content is rendered within the context of a suspending function.
        An example of this approach is the HtmlFlow template engine, which
        supports suspendable templates with the use of the \texttt{Flow} class from the
        Kotlin standard library.
\end{itemize}
% It also includes methods using Kotlin coroutines and other
% asynchronous mechanisms supported by template engines, such as the
% \texttt{writeAsync} method from HtmlFlow, which enables non-blocking I/O
% using \textit{continuations}. 
The Spring MVC implementation uses handlers based solely on the blocking
interface of the \texttt{Observable} class. To enable PSSR in this context, we
utilize the \texttt{StreamingResponseBody} interface, which allows the
application to write directly to the response \texttt{OutputStream} without
blocking the servlet container thread. According to the Spring documentation,
this class is \textit{a controller method return value type for asynchronous
request processing where the application can write directly to the response
OutputStream without holding up the Servlet container thread.}

@@

This implementation includes two main approaches to PSSR\@:
\begin{itemize}
      \item \textbf{Blocking}: The template engine is used in a blocking context,
            where the HTML content is rendered using the blocking interface of the
            \texttt{Observable} class.
      \item \textbf{Virtual}: The template engine is used in a non-blocking context,
            where the HTML content is rendered within the context of Virtual Threads.
\end{itemize}

@@

This implementation includes three main approaches to PSSR\@:
\begin{itemize}
      \item \textbf{Blocking}: The template engine is used in a blocking context,
            where the HTML content is rendered using the blocking interface of the
            \texttt{Observable} class.
      \item \textbf{Virtual}: The template engine is used in a non-blocking context,
            where the HTML content is rendered within the context of Virtual Threads.
      \item \textbf{Reactive}: The template engine is used in a reactive context,
            where the HTML content is rendered using the reactive programming model.
            In this case, we use the HtmlFlow template engine, which
            supports asynchronous rendering through the \texttt{writeAsync} method.
\end{itemize}
